---
title: "PISA2012_wenchit"
output: pdf_document
---
## Abstract

In this research, I want to explore the education resource distribution in Australia. For insatnce, comparing the education resources between remote and city areas and the discrepancies between private or public schools. Based on the PISA2012 mathematics literacy survey data. Also, I will break down the the relationship between school location and how it relates to the scores in math literacy, the relationship between private or public schools and how it relates to the students' scores in math literacy.

# Introduction
- PISA2012: the term PISA stands for Programme for International Student Assessment is a survey that randomly select 15 year-old students(students who are attending secondary schools) as samples for assessment. In 2012, a total of 65 OECD countries and economics and about half a million 15 year-old students participated in the PISA assessment. Generally, the assessed reaults lies in 5 categories: Level 5 are high performers, whereas students who lies below the international standard baseline level 2 are cosidered low performers. In 2012, 775 Australian schools and 14,481 studnets participated in this assessment. To ensure the authenticity, an amount of indigenous students were also sampled. 
Students are categorized in 6 different levels in accordance to their proficiency in the assessment(PISA, 2012):
Level 6: can conceptualise, generalise and use information. They are capable of advanced mathematical thinking and reasoning; have a mastery of symbolic and formal mathematical operations and relationships; and can formulate and precisely communicate their  ndings, interpretations and arguments. Students who score higher than 669.3 scores belongs to level6.
Level 5: can develop and work with models for complex situations; select, compare and evaluate appropriate problem-solving strategies for dealing with complex problems; work strategically using broad, well-developed thinking and reasoning skills; and re ect on their work and formulate and communicate their interpretations and reasoning. Students who score higher than 607.0 scores belongs to level5.
Level 4: can work effectively with explicit models for complex, concrete situations; select and integrate different representations, including symbolic representations; utilise their skills and reason with insight; and construct and communicate explanations and arguments. Students who score higher than 544.7 scores belongs to level4.
Level 3: can execute clearly described procedures, including those that require sequential decisions; select and apply simple problem-solving strategies; interpret and use representations; typically show some ability to handle percentages, fractions and decimal numbers, and to work with proportional relationships; and provide solutions that re ect that they have engaged in basic interpretation and reasoning. Students who score higher than 482.4 scores belongs to level3.
Level 2: can interpret and recognise situations in contexts that require no more than direct inference; extract relevant information from a single source and make use of a single representational mode; employ basic algorithms, formulas, procedures or conventions to solve problems involving whole numbers; and make literal interpretations of the results. Students who score higher than 420.1 scores belongs to level2.
Level 1: can answer questions involving familiar contexts where all relevant information is present and the questions are clearly de ned; identify information and carry out routine procedures; and perform actions that are almost always obvious and follow immediately from the given stimuli. Students who score higher than 357.8 scores belongs to level1.
Below 1: not demonstrate even the most basic types of mathematical literacy that PISA measures. These students are likely to be seriously disadvantaged in their lives beyond school.

- Mathematic Lieteracy: in the mathematic literacy domain, the assessment focused on students' ability to solve mathematic problems described in a real-life situation. In PISA2012 framework(OECD2012, p25), it defined mathematic literacy as follows: "Mathematic literacy is an individual's capacity to formulate, employ and interpret mathematics in a variety of contexts. It includes reasoning mathematically nd using mathematical concepts, procedures, facts and tools to describe, explain and predict phenomena. It assists individuals to recognise the role that Mathematics plays in the world and to make the well-founded judgments and decisions needed by constructive, engaged and reflective citizens."

The assessment in math literacy is designed according three main components:
1. the context of a challenge or problem that arises in the real world
2. the nature of mathematical thought and action that can be used to solve the problem
3. the processes that the problem solver can use to construct a solution.

# Overview of Australian Education System:
--Terms and definition
It is defined in the handbook from PISA that,
Village School or Rural Area: less than 3,000 people 
Small Town School: 3,000 to about 15,000 people
Town School: 15,000 to about 100,000 people
City School: 100,000 to about 1,000,000 people, for example, Hobart, Tasmania
Large City School: ith over 1,000,000 people, for example, Sydney and Melbourne

--Facts
1. In 2016, there are over 9,400 schools in which 1400 of them are secondary schools across Australia.
   Which means more than half of the amount of secondary schools were participants of PISA2012 assessment.
2. In 2016, there are over 6,000 government(public) schools, more than 1700 Catholic schools(private) and
   more than 1,000 independent schools(private)
3. In 2017, around 65% of students attended government school, 19% students attended Catholic schools and
   16% students attended independent schools.

# Methods Used
Linear Mixed Model: including linear regression model and
ANOVA test:



## Results

The following graph shows the amount of private and public schools in Australia that participated:
```{r private/public count plot}
library(data.table)
library(ggplot2)
library(PISA2012lite)
data("school2012")
setDT(school2012)

#plot a bar chart
ggplot(school2012[NC == "Australia" & !is.na(SC01Q01)], aes(x=SC01Q01))+geom_bar(fill="#E69F00", colour="black")+labs(x="Private or Public", y="Total Count")
```


The following is a graph about the distribution of schools in Australia that participated:
```{r school location plot}
library(data.table)
library(ggplot2)
library(PISA2012lite)
data("school2012")
setDT(school2012)

#plot bar chart 
ggplot(school2012[NC == "Australia" & !is.na(SC03Q01)], aes(SC03Q01))+geom_bar(fill="#E69F00", colour="black")+labs(x="School Location", y="Total Count")
```

The following is a calculation of math literacy scores by schools participated all across Australia and summarise by its house weight
and to obtain coefficient in the null model, supposing that school factor is an idependent variable:

```{r null model}
library(data.table)
library(ggplot2)
library(PISA2012lite)
library(dplyr)
library(lme4)
data("computerStudent2012")
setDT(computerStudent2012)

#calculate the weighted math literacy score, using the column W_FSTUWT
n <- nrow(computerStudent2012)
computerStudent2012$W_HOUSEWHT <- n * computerStudent2012$W_FSTUWT / sum(computerStudent2012$W_FSTUWT)

#calculate the weighted average of math literacy scores, using column PV1MATH
computerStudent2012[NC == "Australia"] %>%
  group_by(SCHOOLID) %>%
  summarise(avg1 = weighted.mean(PV1MATH, w = W_HOUSEWHT))

#run linear regression test
HLM0 <- lmer(PV1MATH ~ (1 | SCHOOLID), data = computerStudent2012[NC == "Australia"],
             weights = W_HOUSEWHT)
summary(HLM0)
```
The variance is 

The following cosidering the socio-economic state of the student
```{r ESCS model}
library(data.table)
library(ggplot2)
library(PISA2012lite)
library(dplyr)
library(lme4)
data("computerStudent2012")
setDT(computerStudent2012)

#calculate the weighted math literacy score, using the column W_FSTUWT
n <- nrow(computerStudent2012)
computerStudent2012$W_HOUSEWHT <- n * computerStudent2012$W_FSTUWT / sum(computerStudent2012$W_FSTUWT)

#create a point plot using Australian data
ggplot(data = computerStudent2012[NC == "Australia"], aes(x = ESCS, y = PV1MATH, size = W_HOUSEWHT)) +
  geom_point(aes(colour = SCHOOLID)) + geom_smooth(fill="black", colour="darkblue", size=1)+ theme(legend.position="none")+labs(x="Socio-economic state", y="Math Performance")
```
The graph shows the relationship between socio-economic state of a student and their math performance.
The variance is 


Now we want to look into the relationship between schools and math performances:
```{r school loction math analysis}
library(data.table)
library(PISA2012lite)
library("dplyr")
data("computerStudent2012")
setDT(computerStudent2012)
data("school2012")
setDT(school2012)

#filter out the data from Australia and select column SC03Q01(School location) and SchoolID
school_view <- school2012[NC =="Australia" & !is.na(SC03Q01)]
school_plot_data <- select(school_view, SC03Q01, SCHOOLID)

#calculate the weighted math literacy score, using the column W_FSTUWT and filter data within Australia
n <- nrow(computerStudent2012)
computerStudent2012$W_HOUSEWHT <- n * computerStudent2012$W_FSTUWT / sum(computerStudent2012$W_FSTUWT)
student_data <- computerStudent2012[NC == "Australia" & !is.na(PV1MATH), .(avg= weighted.mean(PV1MATH, w = W_HOUSEWHT)), by=SCHOOLID] 

#merge two tables: student_data and school_plot_data and create a boxplot
plot_data <- merge(student_data, school_plot_data, by="SCHOOLID")
levels <- levels()
ggplot(plot_data, aes(x=factor(SC03Q01), y=avg))+geom_boxplot()+labs(x="School Location", y="Math Performance")

#run linear regression test
fit <- lm(avg~SC03Q01, data=plot_data)
summary(fit)

#run ANOVA test
plot_anova <- aov(avg~SC03Q01, data = plot_data)
summary(plot_anova)

#run Tukey test
tuk <- TukeyHSD(plot_anova)
tuk
plot(tuk)
```
The boxplot shows there is a significant difference in the mean score of students from different region of Australia. 
The graph implies that students from large city in Australia, for example, Sydney and Melbourne normally scores higher than students from a village.

```{r private public analysis}
library(data.table)
library(PISA2012lite)
library("dplyr")
library(ggplot2)
data("computerStudent2012")
setDT(computerStudent2012)
data("school2012")
setDT(school2012)

#filter out the data from Australia and select column SC01Q01(Private or Public School) and SchoolID
school_view <- school2012[NC =="Australia" & !is.na(SC01Q01)]
school_plot_data <- select(school_view, SC01Q01, SCHOOLID)

#calculate the weighted math literacy score, using the column W_FSTUWT and filter data within Australia
n <- nrow(computerStudent2012)
computerStudent2012$W_HOUSEWHT <- n * computerStudent2012$W_FSTUWT / sum(computerStudent2012$W_FSTUWT)
student_data <- computerStudent2012[NC == "Australia" & !is.na(PV1MATH), .(avg= weighted.mean(PV1MATH, w = W_HOUSEWHT)), by=SCHOOLID] 

#merge two tables: student_data and school_plot_data and create a boxplot
plot_data <- merge(student_data, school_plot_data, by="SCHOOLID")
ggplot(plot_data, aes(x=factor(SC01Q01), y=avg))+geom_boxplot()+labs(x="Private or Public", y="Math Performance")

#run linear regression test
fit <- lm(avg~SC01Q01, data=plot_data)
summary(fit)

#run ANOVA test
plot_anova <- aov(avg~SC01Q01, data = plot_data)
summary(plot_anova)
```


## Conclusion
A few results can be concluded accroding to the analysis result:
1. Students from private/independent schools significantly outperform public/government students.
2. Students from large city outperform students from other scetors of students.
3. Students with a higher socio-economic state also show a higher level of proficiency in math literacy
According the Independent Review into Regional, Rural and Remote Education from The Department of Education 
and Training in 2014, the proportion of residents aged between 25~34 years old holding a degree was: Major City 42.2 %, Inner Regional 21.8%, Outer Regional 19.5%, and Remote and Very Remote 17.8%. While there has been growth in the number of people from regional and remote areas 
undertaking an undergraduate degree over the last decade, they remain underrepresented 
in higher education.

## Reference
1. Pmc.gov.au. (2017). Remote School Attendance Strategy | Department of the Prime Minister and Cabinet. [online] Available at: https://www.pmc.gov.au/indigenous-affairs/education/remote-school-attendance-strategy [Accessed 29 Nov. 2017].
2. REGIONAL AND REMOTE EARLY CHILDHOOD EDUCATION. (2017). [pdf] NSW Department of Education. Available at: https://education.nsw.gov.au/media/ecec/pdf-documents/2017/Regional-and-Remote_Literature-Review.pdf [Accessed 30 Nov. 2017].
3. Thomson, S., Bortoli, L. and Buckley, S. (2013). PISA 2012: How Australia measures up. [pdf] Australian Council for Educational Research. Available at: https://research.acer.edu.au/cgi/viewcontent.cgi?article=1015&context=ozpisa [Accessed 2 Dec. 2017].
4. En.wikipedia.org. (2017). Structural equation modeling. [online] Available at: https://en.wikipedia.org/wiki/Structural_equation_modeling [Accessed 4 Dec. 2017].
5. Hox, J. and Bechger, T. (n.d.). An Introduction to Structural Equation Modeling. [ebook] Available at: http://joophox.net/publist/semfamre.pdf [Accessed 4 Dec. 2017].
6. Bulut, O., Delen, E. and Kaya, F. (2012). An SEM Model based on PISA 2009 in Turkey: How Does the use of Technology and Self-regulation Activities Predict Reading Scores?. Procedia - Social and Behavioral Sciences, 64, pp.564-573.
7. Austrade.gov.au. (2017). The Australian education system | Australian Education Technology. [online] Available at: https://www.austrade.gov.au/edtech/the-australian-education-system/ [Accessed 6 Dec. 2017].
8. Halsey, J. (2017). Title: Independent Review into Regional Rural and Remote Education— Discussion Paper. [pdf] The Department of Education and Training. Available at: https://docs.education.gov.au/system/files/doc/other/independent_review_into_regional_rural_and_remote_education.pdf [Accessed 6 Dec. 2017].
